\sysfull

### 2017-11-16 paper/coalescing.tex
coalescing: partial edits to strategies

\textbf{Task Coalescing Strategies.} 
A good coalescing strategy must moderate the risk of wasted work and capitalize
on benefits of deferred commits by adapting the target size of the dynamic task
to the energy available at runtime.
%
The general structure of our strategies is to increase the target task size by
some positive delta $C_\text{up}$ after the dynamic coalesced task completes,
and decrease it by some negative delta $C_\text{down}$ upon a reboot after a
power failure.
%
The choice of $C_\text{up}$ and $C_\text{down}$, including their units, defines
the strategy and its performance.
%
For example, a naive strategy that fixes $C_\text{up}$ and $C_\text{down}$ to
``one static task``, will be very slow to react to changes in incoming energy,
since its dynamic task size can only change by the size of one (small) static
task per charge-discharge cycle.
%
Working from these observations, we propose and evaluate two coalescing
strategies for \sys:

\begin{enumerate}
	\item \textbf{Interrupt History Coalescing (HDC):} when more tasks are coalesced without power interrupt, $C_U$ increases, and $C_D$ increases, as more task are interrupted before committing; and 
	\item \textbf{Weighted Interrupt History Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
\end{enumerate}

\subsection{Power Interrupt History Coalescing}

With this strategy \sys calculates only one parameter $C=C_U=C_D$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period. \sys begins execution in a coalesced task with $S_t=0$. When $S_t=C_{t-1}-1$ coalesced tasks have been executed without power interrupt (where $C_{t-1}$ defines $C$ until $t-1$ task commit) a task commit follows. After each commit, \sys assigns new coalescing factor as $C_{t}=\lceil \sum_{k=1}^{t}S_{k-1}/2\rceil$. This implies that if more tasks complete in an operating period, \sys attempts to coalesce more tasks. However, tasks coalescing are increased by a half the speed of the successful task completion. This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt \sys does not immediately reduce coalescing factor to one, and with successful page commit of coalescing task, \sys gradually increases coalescing factor. In both cases \sys reduces the effect of outliers, i.e. short bursts of energy or short bursts of frequent power interrupts.

@@

\subsection{Weighted Power Interrupt History Coalescing}

The previous coalescing strategy is oblivious to the task size. Although many tasks can be coalesced, their non-equal size will determine that some coalesced tasks will be more susceptible to power interrupts than others. To understand that effect we provide the following illustration.

\textbf{Task Size Profiling.} We have used a set of program benchmarks (refer to Section~\ref{sec:results_software} for details and to the beginning of Section~\ref{sec:methodology_evaluation} for system setup). Each benchmark was run for predefined period of time (in case of Figure~\ref{fig:task_profiling} 10\,seconds), on non-intermittent power, and at each task evocation, its execution time was measured. If a task was evoked more than once, longer execution time replaced the short one---this way we obtained conservative values of task sizes. At the end of task measurement, normalized integer values of task execution for task $x$, $W_x$, are collected---the shortest one being one task, the longest one $W_{x,\max}$ (in case of Figure~\ref{fig:task_profiling} $W_{x,\max}=5$). Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}. We observe huge variations of task sizes---both within a benchmark, as well per benchmark, which calls for the use of this information in coalescing.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}%coalescing_timeline
	\caption{Result of benchmark applications (Section~\ref{sec:results_software}) task size measurement, measured in the number of instruction cycles. We observe very large variations of task sizes per each benchmark. \todo{Describe X axis, no capitals in titles - except first word; markers in addition to colors per app; remove 4.7}{Amjad}}
	\label{fig:task_profiling}
\end{figure}

With the task size information, in the second (weighted) coalescing strategy an application needs to be profiled, as above, before it is loaded to the intermittently-powered device. Then the process of coalescing is the same as for the (non-weighted) power interrupt history coalescing, with the exception that $S_t=W_x \times S_t$, i.e. number of traversed tasks before commit are a function of their size.

>>>>

\textbf{Task Coalescing Strategies.}

The likelihood that a task group will complete depends on the total amount of
energy consumed by all tasks in the group, i.e. the size of the dynamic task.
%
A good coalescing strategy must moderate the risk of wasted work and capitalize
on benefits of deferred commits by adapting the target size of the dynamic task
to the energy available at runtime.
%
The general structure of our strategies is to increase the target task size by
some positive delta $C_\text{up}$ after the dynamic coalesced task completes,
and decrease it by some negative delta $C_\text{down}$ upon a reboot after a
power failure.
%
The choice of $C_\text{up}$ and $C_\text{down}$, including their units, defines
the strategy and its performance.
%
For example, a naive strategy that fixes $C_\text{up}$ and $C_\text{down}$ to
``one static task``, will be very slow to react to changes in incoming energy,
since its dynamic task size can only change by the size of one (small) static
task per charge-discharge cycle.
%
Working from these observations, we propose and evaluate two coalescing
strategies for \sys: \textbf{History-aware Coalescing (HC)} and
\textbf{Task-size- and History-aware Coalescing (TSHC)}.

%\begin{enumerate}
%	\item \textbf{Interrupt History Coalescing (HDC):} when more tasks are coalesced without power interrupt, $C_U$ increases, and $C_D$ increases, as more task are interrupted before committing; and 
%	\item \textbf{Weighted Interrupt History Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
%\end{enumerate}

\subsection{HC: History-aware Coalescing}

\todo{clarify}{AC}
With this strategy \sys calculates only one parameter $C=C_U=C_D$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period. \sys begins execution in a coalesced task with $S_t=0$. When $S_t=C_{t-1}-1$ coalesced tasks have been executed without power interrupt (where $C_{t-1}$ defines $C$ until $t-1$ task commit) a task commit follows. After each commit, \sys assigns new coalescing factor as $C_{t}=\lceil \sum_{k=1}^{t}S_{k-1}/2\rceil$. This implies that if more tasks complete in an operating period, \sys attempts to coalesce more tasks. However, tasks coalescing are increased by a half the speed of the successful task completion. This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt \sys does not immediately reduce coalescing factor to one, and with successful page commit of coalescing task, \sys gradually increases coalescing factor. In both cases \sys reduces the effect of outliers, i.e. short bursts of energy or short bursts of frequent power interrupts.

@@

\subsection{TSHC: Task-size- and History-aware Coalescing}

The history-aware coalescing policy used the number of tasks in the group as a
rough proxy for the energy cost of the task group.
%
With some up-front investment of developer effort and modest time and space
overhead, a more precise proxy for task size can be obtained by profiling the
tasks at design time.
%
The TSHC policy approximates task energy cost from a profile of task execution
times.
%
Profiling execution time instead of direct energy consumption is
advantageous, because the former does not require specialized
current\hyp{}measurement hardware and can be readily done on any
off-the-shelf microcontroller.
%
To use the TSHC policy, the application is first profiled, as will be described
in the experiment below, and the task execution time for each task is written
into a read-only table in non-volatile memory.
%
At runtime, the decision to coalesce or commit at each task boundary proceeds
similarly with the HC policy, but the task completion counter, $S_t$ has units
of ``task execution time''.\todo{be precise}{AC}

To quantify the importance of task size in the coalescing decision, we measured
the distribution of task sizes in our benchmarks (described in
Sections~\ref{sec:results_software}).
%
Each benchmark was run for 10 seconds on non-intermittent power, and a timestamp
in clock cycles was taken at each task boundary. For each execution of
a task, its size was estimated as the maximum difference in timestamps at its
boundaries.
%
The CDF for the distribution of task sizes in Figure~\ref{fig:task_profiling}
shows the wide range of task sizes across and within benchmarks, and explains
the advantage of the TSHC policy.

%At the end of
%task measurement, normalized integer values of task execution for task $x$, $W_x$, are
%collected---the shortest one being one task, the longest one $W_{x,\max}$ (in
%case of Figure~\ref{fig:task_profiling} $W_{x,\max}=5$). Result of the task
%size measurement is provided in Figure~\ref{fig:task_profiling}.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}%coalescing_timeline
	\caption{\textbf{Distribution of task sizes, estimated by task execution times.} The
wide variation in task size motivates accounting for task size in the coalescing strategy.
\todo{Describe X axis, no capitals in titles - except first word; markers in addition to colors per app; remove 4.7}{Amjad}}
	\label{fig:task_profiling}
\end{figure}
